Assessment of drought impact over land cover in continental Chile by the analysis of water supply and demand from ERA5-Land and MODIS

### 2024-01-05 article/paper_drought_Chile2023.qmd
guardar cambios previos

title: Assessment of drought impact over land cover in continental Chile by the analysis of water supply and demand from the ERA5-Land and MODIS

@@

![(a) Chile and the five zones "norte grande", "norte chico",  "zona central", "zona sur", and "zona austral". (b) Topography reference map. (c) Land cover classes for 2021. (d) Persistent land cover classes (> 80%) for 2001-2021.](../output/figs/map_study_con_landcover.png){#fig1_study_area width=80% fig-pos='ht'}

@@

@@

In this study, RL06.1_V3 GRACE mascon (mass concentration) monthly solutions were used, which are provided by the Jet Propulsion Laboratory (JPL-M, https://grace.jpl.nasa.gov). Each GRACE Tellus monthly grid at 0.5 degrees represents the deviation of surface mass for that month relative to a reference time average (2004-2009 baseline), which is subtracted from all other monthly grids to provide terrestrial water storage anomalies (TWSA) [@Ramjeawon2022; @Yin2022]. This JPL-M version of the data employs a Coastal Resolution Improvement (CRI) filter that reduces signal leakage errors across coastlines [@Wiese2019; @Wiese2016]. Although the mascon solutions greatly reduced leakage errors, a gain factor, which is used to enhance the spatial resolution, was applied to the dataset as recommended for hydrological studies [@Ramjeawon2022; @Yin2022]. The water storage and height anomalies are given in Equivalent Water Height or Thickness units (EWH, cm) [@Sabzehee2023], and their temporal resolution monthly is from April 2002 to May 2023.

>>>>

title: Assessment of drought impact over land cover in continental Chile by the analysis of water supply and demand from ERA5-Land and MODIS

@@

![(a) Chile and the five zones "norte grande", "norte chico",  "zona central", "zona sur", and "zona austral". (b) Topography reference map. (c) Land cover classes for 2021. (d) Persistent land cover classes (> 80%) for 2001-2021.](../output/figs/map_study_con_landcover.png){fig-pos='!ht'}

@@

\begin{table}[!ht]
\caption{Value of linear change trend next to time-series plot of surface, per landcover class (IGBP MCD12Q1.006) for 2001-2019 through Central Chile. Red dots on the plots indicate the maximum and minimum surface}
\includegraphics[]{../output/figs/table_var_landcover_macro.png}
\end{table}

@@

In this study, RL06.1_V3 GRACE mascon (mass concentration) monthly solutions were used, which are provided by the Jet Propulsion Laboratory (JPL-M, https://grace.jpl.nasa.gov). Each GRACE Tellus monthly grid at 0.5 degrees represents the deviation of surface mass for that month relative to a reference time average (2004-2009 baseline), which is subtracted from all other monthly grids to provide terrestrial water storage anomalies (TWSA) [@Ramjeawon2022; @Yin2022]. This JPL-M version of the data employs a Coastal Resolution Improvement (CRI) filter that reduces signal leakage errors across coastlines [@NASA2019; @Wiese2016]. Although the mascon solutions greatly reduced leakage errors, a gain factor, which is used to enhance the spatial resolution, was applied to the dataset as recommended for hydrological studies [@Ramjeawon2022; @Yin2022]. The water storage and height anomalies are given in Equivalent Water Height or Thickness units (EWH, cm) [@Sabzehee2023], and their temporal resolution monthly is from April 2002 to May 2023.
